### 2016-12-12 paper/style.tex
Update paper template

%!TEX root=main.tex
\usepackage{tikz}

\usepackage{amssymb}
\usepackage{amsmath}

\usepackage{newfloat}
\DeclareFloatingEnvironment[
  fileext=lol,
  name=Listing
]{listing}

\usepackage[labelformat=simple]{subcaption}

\DeclareCaptionSubType{listing}
\renewcommand\thesublisting{(\alph{sublisting})}

\usepackage{fancybox}

\makeatletter
\newenvironment{CenteredBox}{% 
\begin{Sbox}}{% Save the content in a box
\end{Sbox}\centerline{\parbox{\wd\@Sbox}{\TheSbox}}}% And output it centered
\makeatother

\newcommand{\code}[1]{\texttt{#1}}

\newlength\dualpgfwidth
\setlength\dualpgfwidth{.5\linewidth}

\usepackage{pgfplots}