### 2016-12-12 paper/style.tex
Correct latex warnings

%!TEX root=paper.tex
\usepackage{tikz}

\usepackage{amssymb}
\usepackage{amsmath}

\usepackage{newfloat}
\DeclareFloatingEnvironment[
  fileext=lol,
  name=Listing
]{listing}

\usepackage[labelformat=simple]{subcaption}

\DeclareCaptionSubType{listing}
\renewcommand\thesublisting{(\alph{sublisting})}

\newcommand{\code}[1]{\texttt{#1}}

\newlength\dualpgfwidth
\setlength\dualpgfwidth{.5\linewidth}

\usepackage{pgfplots}
\pgfplotsset{compat=1.14}